Edge sign prediction in general graphs and Correlation Clustering

### 2017-11-28 cc/cc_state_of_art.tex
add some ref
add two refs
tiny fixes and ref all over the places
typos and stability in cc

Because of the complexity of the \pcc{}, one has to rely on approximations to solve large instances
of the problem. However, we can imagine offline signed social networks with only few nodes, in which
case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
to evaluate in practice the quality of heuristic methods, albeit in non asymptotic setting. The
methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
fixed parameter algorithms we describe now only handle complete graphs, while the next set of
methods also work on general graphs.

@@

Let us first describe the cluster editing problem, where one is given a general unsigned input graph
$H=(V,E)$ and wants to
find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
union of cliques. It is possible to navigate between an instance $H$ of cluster editing and an
instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
are deleted from $H$. To the best of our knowledge, the problem was first introduced under this
name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
whereas we shall only give important and more recent pointers here.} They show the problem is
\NPc{}, even if the number of clusters $K\geq2$ is set beforehand\footnote{The reduction is from
3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
$0.878$-approximation in the weighted $K=2$ case using the celebrated Goemans-Williamson SDP
relaxation~\autocite{MaxCutSDP95}.

@@

Although the same problem was considered earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
coin the term \pcc{} and are the first to study this problem complexity. Namely, for complete,
unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
give a $17429$-approximation of \mind{} and a PTAS%
that, for any $\epsilon \in [0,1]$, runs in
$O(n^2e^{O(\epsilon^{-10}\log \epsilon^{-1})})$ and returns with probability
$1-\nicefrac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
solution of \maxa{}.

@@

We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
4.4]{Demaine2006})\Todo{give the intuition of the proof} that it takes a linear time to construct a
multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.

\begin{figure}[htpb]
   \centering
   \includegraphics[width=0.8\linewidth]{assets/raw/cc_to_mmc.png}
   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (temporarily reproduced from
   \autocite{Demaine2006}) \label{fig:cc_mmc}}
\end{figure}

If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
within a factor $c$~\autocite{}. On general graphs, the best known approximation factor in $O(\log
k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
$O(\log|E^-|)=O(\log n^2)=O(\log n)$. This approximation ratio is achieved by first solving the
relaxation of the integer problem \eqref{eq:mindIP}, where the constraint \eqref{eq:mindInteger} is
replaced by $x_{uv}\in [0,1]$. From now on, we refer to that relaxation as the \emph{canonical
\mind{} LP}. Once this canonical \mind{} LP is solved, we interpret $x_{uv}$ as a distance: the
larger it is and the more we want $u$ and $v$ to be in different clusters. We can then use the
\regionGrow{} method~\autocite{RegionGrowing93}. The idea is to pick a random center $u$ and to add
to $u$'s cluster all the nodes at distance less than $r$ from $u$ before removing that cluster and
repeating the process. $r<\shalf$ is chosen adaptively such that the weight of positive edges
leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
inside the cluster.\Todo{The full proof can be found in pages 7--9 of
\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
doesn't have much to do with machine learning.}

@@

Not long after, \textcite{CCPivotConf05} came up with a better approximation. To explain it, we
first describe their randomized combinatorial algorithm \ccpivot{}, which gives a $3$-approximation
of \mind{} on complete unweighted graphs (and was later derandomized while preserving its
approximation guarantee~\autocite{derandomCCPivot08}). At each iteration, we pick a node $u$ \uar{}
(called the pivot) and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the canonical LP to obtain different approximation factor depending on the
constraints imposed on the weight.\Todo{Comment on those constraints, especially the triangular ones
in the context of consensus clustering.} Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, this
tweaking provide a $2.5$-approximation. Note that unweighted graphs naturally fit into that case, as
each edge is either labeled $+$ or $-$. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the canonical LP with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

@@

This concludes the presentation of the known approximation results on \pcc{}, that we gather in
\autoref{tab:cc_approx}.
To summarize, except for maximizing the number of agreements on a complete graphs, we cannot hope
for exact solutions for the other problems in the worst case, since they are \APXh{}. Moreover, even
constant factor approximation are ruled out for \mind{} on general graphs, as showed by the
equivalence with \mmc{}.
Currently there are three main approaches to approximate \mind{}. On a general graph,
one first solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
resulting in an $O(log n)$ approximation. On complete graphs, the simplest solution is the
\ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
the proof of \autoref{thm:weakbal-consistent}. The approximation factor of this general idea can be
improved by interpreting the fractional solution of the linear program as probability to be included
in the pivot's cluster. The cost of solving linear programs with so many variables and constraints
being prohibitive, real world general graphs requires to move from methods with guarantees to more
heuristic approaches we describe next. Before that, we make two remarks. The first one regards a
line of work improving scalability of these approximations by presenting distributed solutions. The
second one is concerned with the case where the number of clusters is set beforehand, and proves
that the problem becomes easier.

@@

While all the methods we discuss so far
are either exact or come with some approximation guarantees, practitioners have also develop
approaches that are designed to efficiently reach a solution that is satisfactory enough for the
application at hand.

@@

After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
\textcite{Elsner2009} describe three folklore heuristics that start with empty clusters and add node
one by one: \enquote{The \textsc{Best} algorithm adds each node $u$ to the cluster with the
strongest $w^\pm$ connecting to $u$, or to a new singleton if none of the $w^\pm$ are positive.
The \textsc{First} algorithm adds each node $u$ to the cluster containing the most recently
considered node $v$ with $w^\pm_{uv} > 0$. The \textsc{Vote} algorithm adds each node to the cluster
that minimizes the \pcc{} objective, \ie{} to the cluster maximizing the total net weight
or to a singleton if no total is positive.} Empirically, \textsc{Vote} turns out to be the best.
Among other related heuristics, \textcite{mergingHeuristics14} describe the random maximum merging
algorithm, that starts with singleton clusters and keep merging two clusters chosen at random among
those whose merge would result in the maximum improvement of the score function. This runs in
$O(n^2\log n)$, and empirically results in fewer disagreements than \ccpivot{}.
Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
greedily build a clustering in fashion similar to \textsc{Vote}, albeit with more randomness in the
node ordering. This clustering is locally improved by moving blocks of $r\in\{1,2\}$ nodes from one
cluster to another as long as the number of disagreements decreases, a phase called neighborhood
descent. The algorithm then enters an inner loop where the current clustering is perturbed by $t$
random one-node-moves and updated if a subsequent neighborhood descent can improve it compared with
before the perturbation. The authors note that both the outer loop and the neighborhood descent can
be run in parallel, which allow them to process a \np{10000} nodes graph on 10 cores in around 700
seconds. \Textcite{heuristicCE16} show how to use ILS to generate initial solutions of the Cluster
Editing problem that are then fed to an integer program.
This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!} Instead of
moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
graph, $G^+$ is a disjoint union of cliques in which all the nodes of a given cluster share the shame
neighbors. Their algorithm is initialized with the set $E_s^+$ of edges where $w^+_{uv} > w^-_{uv}$,
and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the
canonical LP to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
$L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
otherwise~\autocite{AltOptimLP13}. Indeed, $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
is the cluster assignment. By relaxing the integer constraints on $L$, it is possible to do
alternate optimization on $L$ and $\cluster$.

@@

In the case of $2$-\pcc{}, \textcite{Facchetti2011isingmodel} rewrite the Hamiltonian as
$\mathcal{H}(\bm{\sigma}) = -\frac{1}{2}\bm{\sigma}^T A \bm{\sigma} = -\frac{1}{2}\bm{1}^TT_\sigma A
T_\sigma\bm{1}$. There $T_\sigma = diag(\bm{\sigma})$ (where $\bm{\sigma} \in \{0,1\}^n$) is the
outcome of a local search algorithm such that $A_\sigma = T_\sigma A T_\sigma$ has the same number
of disagreements as $A$ but the smallest number of negative edges. This is called a gauge
transformation in the spin glass literature and the benefit of that heuristic is that it scales
gracefully to large graphs.
\Textcite{Bagon2011} also write the \maxa{} objective as a Potts model, and show that it can be
interpreted as the log posterior of a partition matrix under a simple generative model and as a
pair-wise conditional random field energy without unary terms. This allows them to adapt existing
discrete energy optimization algorithms in order to cope with the following three challenges of the
\pcc{} energy: \enquote{(i) the energy is non sub-modular, (ii) the number of clusters is not known
in advance, and (iii) there is no unary term}. Doing so, they are able to handle large problems with
more than 100K nodes.  Also adopting an energy minimization approach, \textcite{Kappes2016} assign a
probability to each cut of a signed graph proportional to the exponential of the number of
disagreements of that cut. They also develop efficient cut sampling methods.

@@

Case in point, in the context of entity resolution, \textcite{activeCoref07} consider the \pcc{}
problem where, by querying an oracle, one can either reduce the uncertainty about one edge weight or
add an extra node with edges connecting it to existing nodes. However, this requires web queries
that are resource bounded and thus yields an active setting learning problem, where one has to
choose the most informative
queries given a budget. A formal definition is given in~\autocite{queryMoreEdgeCC07} and in
practice, after finding an initially good partition, they select in each cluster a node to be its
centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
In the case the oracle answering the queries is consistent, then there exists an information
theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
that bound up to a $O(\log n)$ factor~\autocite{perfectActiveLong17}.
\textcite[Section 5]{Ailon2014} present another active algorithm that solves \mind$[k]$ with a
linear number of queries but an exponential running time in $n$.  In a noisy setting,
\textcite{Mitzenmacher2016} assume there is a planted $k$-partition and that we have access to an
oracle that for $u,v\in V^2$ returns $\cluster(u) - \cluster(v) \mod k$ with probability $1-p$ and a
noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{},
whereas for $k>2$, the oracle gives more information than simply the sign of the path from $u$ to
$v$. Whenever
$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
the planted partition in polynomial time and with high probability. Those queries are actually
random (\ie{} non-adaptive), and the clusters are found by looking at almost edge-disjoint path
between all pairs of nodes.

@@

This state of the art depicts a nuanced landscape for the \pcc{} problem. On one hand, minimizing
the disagreements is difficult, although it is the most natural way given our bias. On the other
hand, it can be solved with approximation guarantees, at least in theory. Indeed, the best methods
requires solving linear programs with $m$ variables and up to $O(n^3)$ constraints, which results in
a high complexity of $O(m^{4.5})$~\autocite[Section 7.2]{LPBook07}. Fortunately we can obtain
reasonable solutions in practice, thanks to efficient heuristics (using the output of \ccpivot{} as
initialization for instance). Furthermore, instances obeying our bias can be solved more easily and
almost exactly. We next present settings where instances are indeed close to our bias.
Before that, we remark that whereas \ccpivot{} is often used on general graphs in practice by
assuming the missing edges are negative, doing so does not preserve its approximation guarantee.
\begin{marginfigure}
  \centering
  \includegraphics[width=\linewidth]{assets/raw/pivotstar.pdf}
  \caption{A positive star with few negative edges}
  \label{fig:cc_star_pivot}
\end{marginfigure}

>>>>

Because of the complexity of the \pcc{}, one has to rely on approximations to solve large instances
of the problem. However, we can imagine offline signed social networks with only few nodes, in which
case it is reasonable to expect finding the optimal clustering. Furthermore, this can also be useful
to evaluate in practice the quality of heuristic methods, albeit in non asymptotic settings. The
methods used to solve \pcc{} exactly depends on the type of graph considered. More precisely, the
fixed parameter algorithms we describe now only handle complete graphs, while the next set of
methods also work on general graphs.

@@

Let us first describe the cluster editing problem, where one is given a general unsigned input graph
$H=(V,E)$ and wants to
find the smallest number of edges that have to be added or deleted to turn $H$ into node-disjoint
union of cliques. It is possible to alternate between an instance $H$ of cluster editing and an
instance $G$ of \pcc{} on complete graphs in polynomial time. To see why, let $G=(V,(E^+,E^-))$
be a complete signed graph with nodeset $V$, $E^+$ be the set $E$ of edges of $H$ and $E^-$ be all
the edges that are not in $H$. The optimal clusters of $G$ are the cliques of $H$, the negative
disagreements within clusters are edges added to $H$ and the positive disagreements between clusters
are deleted from $H$. To the best of our knowledge, the problem was first introduced under this
name by \textcite{Shamir02}.\marginpars{However, the same problem was studied before and we refer the
reader to the comprehensive survey of \textcite{clusterEditSurvey13} for additional details,
whereas we shall only give important and more recent pointers here.} They show the problem is
\NPc{}, even if the number of clusters $K\geq2$ is set beforehand\footnote{The reduction is from
3-exact 3 cover, see \autocite[Theorems 1, 2 and Corollary 1]{Shamir02}.} and provide a
$0.878$-approximation in the weighted $K=2$ case using the celebrated Goemans-Williamson SDP
relaxation~\autocite{MaxCutSDP95}.

@@

Although the same problem was considered earlier~\autocites{Early96}{Ben-Dor99}, \textcite{Bansal2002}
coin the term \pcc{} and are the first to study this problem complexity. Namely, for complete,
unweighted signed graphs, they show that both \mind{} and \maxa{} are \NPc{}. Along the way, they
give a $17429$-approximation of \mind{} and a PTAS
that, for any $\epsilon \in [0,1]$, runs in
$O(n^2e^{O(\epsilon^{-10}\log \epsilon^{-1})})$ and returns with probability
$1-\nicefrac{\epsilon}{3}$ a solution with at most $\epsilon n^2$ fewer agreements than the optimal
solution of \maxa{}.

@@

We next present the polynomial transformation from \pcc{} to \mmc{}. Given $G=(V, (E^+, E^-), w)$ we let
$H_G$ be the graph induced by $E^+$ with the same weight. Then for every negative edge $(u,v)$ of
weight $w_{uv}^-$, we create a new vertex $v_{\widehat{uv}}$, connect $u$ to $v_{\widehat{uv}}$ with
weight $w_{uv}^-$ and let $(v_{\widehat{uv}}, v)$ be a source-target pair added to $S_G$. This
construction is depicted in \autoref{fig:cc_mmc} and one can show (\autocite[Theorem
4.4]{Demaine2006}) that it takes a linear time to construct a
multicut of weight $W$ in $H_G$ from a clustering of weight $W$ in $G$, and vice versa.
%\Todo{give the intuition of the proof}

\begin{figure}[htpb]
   \centering
   \includegraphics[width=0.8\linewidth]{assets/raw/cc_to_mmc.png}
   \caption{The transformation from \pcc{} on $G$ to \mmc{} on $H$ (reproduced from
   \autocite{Demaine2006}) \label{fig:cc_mmc}}
\end{figure}

If a certain conjecture in computational complexity is true\footnote{Namely, the Unique Game
Conjecture of~\textcite{UGC02}.}, then for every constant $c>0$ it is \NPh{} to approximate \mmc{}
within a factor $c$~\autocite{MulticutHardness06}. On general graphs, the best known approximation factor in $O(\log
k)$~\autocite{RegionGrowing93}, which according to the presented reduction translates to
$O(\log|E^-|)=O(\log n^2)=O(\log n)$. This approximation ratio is achieved by first solving the
relaxation of the integer problem \eqref{eq:mindIP}, where the constraint \eqref{eq:mindInteger} is
replaced by $x_{uv}\in [0,1]$. From now on, we refer to that relaxation as the \emph{canonical
\mind{} LP}. Once this canonical \mind{} LP is solved, we interpret $x_{uv}$ as a distance: the
larger it is and the more we want $u$ and $v$ to be in different clusters. We can then use the
\regionGrow{} method~\autocite{RegionGrowing93}. The idea is to pick a random center $u$ and to add
to $u$'s cluster all the nodes at distance less than $r$ from $u$ before removing that cluster and
repeating the process. $r<\shalf$ is chosen adaptively such that the weight of positive edges
leaving the cluster is less than $c\log(n+1)$ times the fractional weights of the positive edges
inside the cluster.
%\Todo{The full proof can be found in pages 7--9 of
%\href{http://erikdemaine.org/papers/Clustering_TCS/paper.pdf}{this online pdf} but in my opinion, it
%doesn't have much to do with machine learning.}

@@

Not long after, \textcite{CCPivotConf05} came up with a better approximation. To explain it, we
first describe their randomized combinatorial algorithm \ccpivot{}, which gives a $3$-approximation
of \mind{} on complete unweighted graphs (and was later derandomized while preserving its
approximation guarantee~\autocite{derandomCCPivot08}). At each iteration, we pick a node $u$ \uar{}
(called the pivot) and we create a cluster containing $u$ and all its neighbors
linked by a positive edges. On weighted complete graphs, they tweak this algorithm by using the
solution of the canonical LP to obtain different approximation factor depending on the
% \Todo{Comment on those constraints, especially the triangular ones in the context of consensus clustering.}
constraints imposed on the weight. Recall that in the general formulation of the problem, each
edge carries two positive numbers: $w_{u,v}^+$ and $w_{u,v}^-$. If the weights respect the
probability constraints stating that for all edge $(u,v)$ in $E$, $w_{u,v}^+ + w_{u,v}^- = 1$, this
tweaking provide a $2.5$-approximation. Note that unweighted graphs naturally fit into that case, as
each edge is either labeled $+$ or $-$. If the weights
additionally respect the triangular inequality constraints stating that $w_{u,v}^- \leq w_{u,w}^- +
w_{w,v}^-$, this become a $2$-approximation. After solving the canonical LP with additional
probability constraints, when picking a node $u$, each of its neighbors $v$ is added to the cluster
of $u$ with probability $x_{uv}$. \Textcite{Chawla2014} improve these two factors to respectively
$2.06$ and $1.5$ by exploiting the same idea but setting the probability to include each neighbor
$v$ of $u$ in the cluster of $u$ to be $1-f^+(x_{uv})$ if $(u,v)\in E^+$ and $1-f^-(x_{uv})$ if
$(u,v)\in E^-$, with a careful choice of $f^+$ and $f^-$. They also give a derandomized version of
their algorithm in the full version of the paper~\autocite[Theorem 23]{ChawlaArxiv14}.

@@

This concludes the presentation of the known approximation results on \pcc{}, that we gather in
\autoref{tab:cc_approx}.
To summarize, except for maximizing the number of agreements on a complete graphs, we cannot hope
for exact solutions for the other problems in the worst case, since they are \APXh{}. Moreover, even
constant factor approximation are ruled out for \mind{} on general graphs, as showed by the
equivalence with \mmc{}.
Currently there are three main approaches to approximate \mind{}. On a general graph,
one first solve the canonical LP, and use the \regionGrow{} method to round the fractional solution,
resulting in an $O(\log n)$ approximation. On complete graphs, the simplest solution is the
\ccpivot{} algorithm, that grows clusters out of randomly selected pivots, in a way reminiscent of
the proof of \autoref{thm:weakbal-consistent}. The approximation factor of this general idea can be
improved by interpreting the fractional solution of the linear program as probability to be included
in the pivot's cluster. The cost of solving linear programs with so many variables and constraints
being prohibitive, real world general graphs requires to move from methods with guarantees to more
heuristic approaches we describe next. Before that, we make two remarks. The first one regards a
line of work improving scalability of these approximations by presenting distributed solutions. The
second one is concerned with the case where the number of clusters is set beforehand, and proves
that the problem becomes easier.

@@

While all the methods we discuss so far
are either exact or come with some approximation guarantees, practitioners have also develop
approaches that are designed to efficiently reach a solution that is satisfactory enough for the
application at hand.
% \url{http://ieeexplore.ieee.org/document/7321476}

@@

After defining the net weight of an edge to be $w^\pm_{uv} = w^+_{uv} - w^-_{uv}$,
\textcite{Elsner2009} describe three folklore heuristics that start with empty clusters and add node
one by one: \enquote{The \textsc{Best} algorithm adds each node $u$ to the cluster with the
strongest $w^\pm$ connecting to $u$, or to a new singleton if none of the $w^\pm$ are positive.
The \textsc{First} algorithm adds each node $u$ to the cluster containing the most recently
considered node $v$ with $w^\pm_{uv} > 0$. The \textsc{Vote} algorithm adds each node to the cluster
that minimizes the \pcc{} objective, \ie{} to the cluster maximizing the total net weight
or to a singleton if no total is positive.} Empirically, \textsc{Vote} turns out to be the best.
Among other related heuristics, \textcite{mergingHeuristics14} describe the random maximum merging
algorithm, that starts with singleton clusters and keep merging two clusters chosen at random among
those whose merge would result in the maximum improvement of the score function. This runs in
$O(n^2\log n)$, and empirically results in fewer disagreements than \ccpivot{}.
Building upon their previous GRASP work~\autocite{GRASP13}, an iterated local search (ILS) heuristic is
presented in~\autocites{Levorato2015}{Levorato2017}. Each iteration of this algorithm starts by
greedily build a clustering in fashion similar to \textsc{Vote}, albeit with more randomness in the
node ordering. This clustering is locally improved by moving blocks of $r\in\{1,2\}$ nodes from one
cluster to another as long as the number of disagreements decreases, a phase called neighborhood
descent. The algorithm then enters an inner loop where the current clustering is perturbed by $t$
random one-node-moves and updated if a subsequent neighborhood descent can improve it compared with
before the perturbation. The authors note that both the outer loop and the neighborhood descent can
be run in parallel, which allow them to process a \np{10000} nodes graph on 10 cores in around 700
seconds. \Textcite{heuristicCE16} show how to use ILS to generate initial solutions of the Cluster
Editing problem that are then fed to an integer program.
% This local search followed by random perturbation is reminiscent of genetic algorithms, and such an
% approach is sketched in~\autocite{GeneticCC08}.\footnote{but that's a terrible paper!}
Instead of
moving nodes, \textcite{restoreNeighborhood13} starts from the observation that in a balanced
graph, $G^+$ is a disjoint union of cliques in which all the nodes of a given cluster share the shame
neighbors. Their algorithm is initialized with the set $E_s^+$ of edges where $w^+_{uv} > w^-_{uv}$,
and repeatedly samples an edge $(u,v)$ from $E_s^+$ before trying to make the neighborhoods of $u$
and $v$ coincide by adding or removing edges in $E_s^+$.  Yet another idea is to modify the
canonical LP to replace the binary variable $x_{uv}$ by a $k$ clusters indicator matrix
$L\in\Rbb^{k\times n}$ where $L_{iu}$ is $1$ is $u\in C_i$ and $-1$
otherwise~\autocite{AltOptimLP13}. Indeed, $x_{u\cdot} \equiv L_{\cluster(u)\cdot}$, where $\cluster$
is the cluster assignment. By relaxing the integer constraints on $L$, it is possible to do
alternate optimization on $L$ and $\cluster$.

@@

In the case of $2$-\pcc{}, \textcite{Facchetti2011isingmodel} rewrite the Hamiltonian as
$\mathcal{H}(\bm{\sigma}) = -\frac{1}{2}\bm{\sigma}^T A \bm{\sigma} = -\frac{1}{2}\bm{1}^TT_\sigma A
T_\sigma\bm{1}$. There $T_\sigma = diag(\bm{\sigma})$ (where $\bm{\sigma} \in \{0,1\}^n$) is the
outcome of a local search algorithm such that $A_\sigma = T_\sigma A T_\sigma$ has the same number
of disagreements as $A$ but the smallest number of negative edges. This is called a gauge
transformation in the spin glass literature and the benefit of that heuristic is that it scales
gracefully to large graphs.
\Textcite{Bagon2011} also write the \maxa{} objective as a Potts model, and show that it can be
interpreted as the log posterior of a partition matrix under a simple generative model and as a
pair-wise conditional random field energy without unary terms. This allows them to adapt existing
discrete energy optimization algorithms in order to cope with the following three challenges of the
\pcc{} energy: \enquote{(i) the energy is non sub-modular, (ii) the number of clusters is not known
in advance, and (iii) there is no unary term}. Doing so, they are able to handle large problems with
more than 100K nodes.  Also adopting an energy minimization approach, \textcite{Kappes2016} assign a
probability to each cut of a signed graph proportional to the exponential of the number of
disagreements of that cut. They also develop efficient cut sampling methods. Several of these
methods have recently been evaluated empirically by \textcite{energyCCBench17}.

@@

Case in point, in the context of entity resolution, \textcite{activeCoref07} consider the \pcc{}
problem where, by querying an oracle, one can either reduce the uncertainty about one edge weight or
add an extra node with edges connecting it to existing nodes. However, this requires web queries
that are resource bounded and thus yields an active setting learning problem, where one has to
choose the most informative
queries given a budget. A formal definition is given in~\autocite{queryMoreEdgeCC07} and in
practice, after finding an initially good partition, they select in each cluster a node to be its
centroid, and query the edges connecting the centroid as ordered by an entropy-based criterion.
In the case the oracle answering the queries is consistent, then there exists an information
theoretic bound on the number of queries needed to recover the clusters, and an algorithm matching
that bound up to a $O(\log n)$ factor~\autocite{perfectActiveLong17}.
\textcite[Section 5]{Ailon2014} present another active algorithm that solves \mind$[k]$ with a
linear number of queries but an exponential running time in $n$.  In a noisy setting,
\textcite{Mitzenmacher2016} assume there is a planted $k$-partition and that we have access to an
oracle that for $u,v\in V^2$ returns $\cluster(u) - \cluster(v) \mod k$ with probability $1-p$ and a
noisy $\cluster(u) - \cluster(v) \pm 1 \mod k$ otherwise. For $k=2$, this is exactly $2$-\pcc{},
whereas for $k>2$, the oracle gives more information than simply the sign of the path from $u$ to
$v$. Whenever
$p < \shalf$, they show that $O(n^{1+\frac{1}{\log\log n}}\log n)$ queries are enough are to recover
the planted partition in polynomial time and with high probability. Those queries are actually
random (\ie{} non-adaptive), and the clusters are found by looking at almost edge-disjoint path
between all pairs of nodes.
% this was improved in https://arxiv.org/abs/1709.07308

@@

This state of the art depicts a nuanced landscape for the \pcc{} problem. On one hand, minimizing
the disagreements is difficult, although it is the most natural way given our bias. On the other
hand, it can be solved with approximation guarantees, at least in theory. Indeed, the best methods
requires solving linear programs with $m$ variables and up to $O(n^3)$ constraints, which results in
a high complexity of $O(m^{4.5})$~\autocite[Section 7.2]{LPBook07}. Fortunately, we can obtain
reasonable solutions in practice, thanks to efficient heuristics (using the output of \ccpivot{} as
initialization for instance). Furthermore, instances obeying our bias can be solved more easily and
almost exactly. We next present settings where instances are indeed close to our bias.
Before that, we remark that whereas \ccpivot{} is often used on general graphs in practice by
assuming the missing edges are negative, doing so does not preserve its approximation guarantee.
\begin{marginfigure}
  \centering
  \includegraphics[width=\linewidth]{assets/raw/pivotstar.pdf}
  \caption{A positive star with few negative edges}
  \label{fig:cc_star_pivot}
\end{marginfigure}
